


Load libraries

```{r}
# https://github.com/etam4260/kneedle
source("Utilities.R")
library(RColorBrewer)
library(dbscan)
library(ggplot2)
library(grid)
library(data.table)
```



Set folder and type of selection function here

```{r}
base_path = "../scripts/example_data/"
sub_folder = "2024-03-12-175828/"
selection_func = "r"

figure_folder = paste0(base_path, sub_folder, "figures/")
dir.create(figure_folder, showWarnings = FALSE)
```


Read data

```{r}
data = read_sim_data(base_path, sub_folder)

df_model_sweep = data[[4]]
df_parameters = data[[1]]
df_planets_sweep = data[[2]]
df_life_sweep = data[[3]]
diffs = df_parameters$r
param_order = "ool r mut id"
selection_func = "r"

# Remove some of the redundant r values
# 12, 11 and 10 have ~ the same values as 20; remove all and keep 20
df_parameters$r = c(9, 20, 50, 174)
df_model_sweep = subset(df_model_sweep, r %in% df_parameters$r)
```


Subset 1 time step of 1 run.

```{r}
n_ool = 1
p_r = 20
p_mutation = 0
p_id = 1

result = select_parameter_set(df_planets_sweep, df_life_sweep, p_id, p_r, p_mutation, selection_func)

df_planets = result[[1]]
df_lifes = result[[2]]
df_model = result[[3]]
```



Time-independent parameters
```{r}
total_n_planets = 1000
select_cols = c(12:21) 
db_minPts = 11
elbow_sens = 1 # empirically determined to correspond to elbow visually
time_steps = unique(df_planets$step)
n_time_steps = length(time_steps)
df_planets_t0 = subset(df_planets, step==0)

# when saving figure
param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)

```



Plot terraformed ratio = f(timestep) for reference

```{r}
t_ratio = data.frame(unique(df_model$step), unique(df_model$n_living_planets)/total_n_planets) 
colnames(t_ratio) = c("step", "alive_ratio")
plot(t_ratio$step, t_ratio$alive_ratio)
```


Cluster compositions through time, or read file if it already exists.
Calculate IQR, mantel, and p-value for each cluster as well.
Time consuming function.

```{r}
run_name = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation, "-elbowsens_", elbow_sens)
clusters_path = paste0(base_path, sub_folder, run_name, "-clusters.csv")
spread_path = paste0(base_path, sub_folder, run_name,"-spread.csv")

max_timestep = max(time_steps)

if(file.exists(clusters_path)){
  print("Clusters file already exist")
  print(clusters_path)
  dbscan_sweep = read.csv(clusters_path)
  print("coefficients read.")
} else{
  sp_spread_columns = c("step", "cluster_id", "cluster_size", "iqr", "mantel", "pval")
  sp_spread_sweep = data.frame(matrix(0, nrow = 0, ncol = length(sp_spread_columns)))
  colnames(sp_spread_sweep) = sp_spread_columns
  
  columns = c("step","planet_id","cluster_id")
  dbscan_sweep = data.frame(matrix(0, nrow = total_n_planets*n_time_steps, ncol = length(columns)))
  colnames(dbscan_sweep) = columns
  
  dbscan_sweep = data.frame(matrix(0, nrow = total_n_planets*n_time_steps, ncol = length(columns)))
  colnames(dbscan_sweep) = columns
  
  for(i in c(1:n_time_steps)){ 
  
    t = time_steps[i]
    print(paste(t, "/", max_timestep, "steps"))
    df_planets_t = subset(df_planets, step == t)
    compositions = df_planets_t[,select_cols]
    comp_distances = full(distance(compositions))
    
    dbscan_cl = dbscan_clusters(db_minPts, comp_distances, total_n_planets, sensitivity = elbow_sens)
    start = (i-1)*total_n_planets+1 # start at 1
    dbscan_sweep$step[start:(start+total_n_planets-1)] = t
    # ids start at 1
    dbscan_sweep$planet_id[start:(start+total_n_planets-1)] = c(1:total_n_planets)
    dbscan_sweep$cluster_id[start:(start+total_n_planets-1)] = dbscan_cl$cluster
    
   
    cids = unique(dbscan_cl$cluster)
    for (cid in cids){
      detected = df_planets_t[dbscan_cl$cluster==cid,]
      # Calculate the interquartile range
      iqr = (IQR(detected$x) + IQR(detected$y) + IQR(detected$z))/3
      
      # Calculate the mantel
      dist_ph_sc = distance(detected[,c("x", "y", "z")])
      dist_comp_sc= distance(detected[select_cols])
      # we should not calculate this here, but instead subset the distance from a global distance matrix
      a = lower(dist_ph_sc)
      b = lower(dist_comp_sc)
      m = mantel(a ~ b, nperm=100)
      row = c(t, cid, length(detected$id), iqr, m["mantelr"], m["pval3"])
      sp_spread_sweep = rbind(sp_spread_sweep, row)
      colnames(sp_spread_sweep) = sp_spread_columns
    }
  }
  
  sp_spread_sweep$alive_ratio = 0
  for(t in sp_spread_sweep$step){
    a = subset(t_ratio, step==t)
    sp_spread_sweep[sp_spread_sweep$step==t, ]$alive_ratio = a$alive_ratio
  }
  
  print(paste0("saving at: ", clusters_path))
  write.csv(dbscan_sweep, clusters_path)
  print(paste0("saving at: ", spread_path))
  write.csv(sp_spread_sweep, spread_path)
}

```


Interquartile ratio threshold

```{r}
# 1/8 of the space in 3D
half_space = subset(df_planets_t0, (x<max(x/2)) & (y<max(y/2)) & (z<max(z/2)))
iqr_threshold = (IQR(half_space$x) + IQR(half_space$y) + IQR(half_space$z))/3
```


Plot the mantel and p-value for each cluster

```{r}
save_pdf = FALSE

if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-clusters_mantel_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 5,
    height = 4)
}

not_noise = subset(sp_spread_sweep, cluster_id != 0)

plot(sp_spread_sweep$alive_ratio, sp_spread_sweep$mantel,
     xlab = "Ratio of terraformed planets",
     ylab = "Mantel coefficient",
     pch=20, col="darkgray")
points(not_noise$alive_ratio, not_noise$mantel, pch=20)
points(df_model$n_living_planets/1000, df_model$mantel_corr_coeff, type="l")
abline(v=0.04, lty=2)

if(save_pdf){
  dev.off()
}

if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-clusters_pvalue_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 5,
    height = 4)
}

plot(sp_spread_sweep$alive_ratio, sp_spread_sweep$pval,
     xlab = "Ratio of terraformed planets",
     ylab = "Mantel p-value",
     pch=20, col="darkgray")
points(not_noise$alive_ratio, not_noise$pval, pch=20)
points(df_model$n_living_planets/1000, df_model$mantel_p_value, type="l")
abline(v=0.04, lty=2)

if(save_pdf){
  dev.off()
}


```

Remove each cluster from the dataset, and calculate the mantel of the remaining space.
Then calculate the mantel contribution of each cluster.

```{r}

if (!"cropped_mantel" %in% colnames(sp_spread_sweep)){
  
  sp_spread_sweep$cropped_mantel = -1
  sp_spread_sweep$cropped_pvalue = -1

  select_cols = c(12:21) 
  print(colnames(df_planets_t)[select_cols])

  for (t in unique(sp_spread_sweep$step)){
    print(paste(t, "/", max_timestep, "steps"))
    
    df_planets_t = subset(df_planets, step == t)
    dbscan_cl = subset(dbscan_sweep, step == t)
    alive_ratio = subset(t_ratio, step == t)$alive_ratio[1]
  
    compositions = df_planets_t[,select_cols]
    comp_distances = full(distance(compositions))

    cids = unique(dbscan_cl$cluster_id)
    for (cid in cids){
      cropped_planets = df_planets_t[!dbscan_cl$cluster_id==cid,]
      dist_ph = distance(cropped_planets[,c("x", "y", "z")])
      dist_comp = distance(cropped_planets[select_cols])
      # we should not calculate this here
      a = lower(dist_ph)
      b = lower(dist_comp)
      m = mantel(a ~ b, nperm=100)
      
      
      sp_spread_sweep[sp_spread_sweep$step==t & sp_spread_sweep$cluster_id==cid,]$cropped_mantel = m["mantelr"]
      sp_spread_sweep[sp_spread_sweep$step==t & sp_spread_sweep$cluster_id==cid,]$cropped_pvalue = m["pval3"]
    }
  }
  
  print(paste0("saving at: ", spread_path))
  write.csv(sp_spread_sweep, spread_path)
}
```


Plot results

```{r}
plot(sp_spread_sweep$alive_ratio, sp_spread_sweep$cropped_mantel,
     ylab = "Cropped space mantel", xlab = "Terraformed ratio"
     )
abline(v=0.04)
plot(sp_spread_sweep$alive_ratio, sp_spread_sweep$cropped_pvalue,
     ylab = "Cropped space p-value", xlab = "Terraformed ratio"
     )
abline(v=0.04)
```


Calculate the contribution of each cluster to the overall mantel

```{r}

# sp_spread_sweep$cropped_mantel = cropped_space_mantels$mantel
sp_spread_sweep$mantel_contribution = -1

for (t in unique(sp_spread_sweep$step)){
  df_model_t = subset(df_model, step == t)
  global_mantel = df_model_t$mantel_corr_coeff
  # difference / global
  # eg: m was 1, now is 0.8; contribution is 1-0.8, normalized
  contribution = (global_mantel - sp_spread_sweep[sp_spread_sweep$step==t,]$cropped_mantel)/global_mantel
  sp_spread_sweep[sp_spread_sweep$step==t,]$mantel_contribution = contribution
}

```

```{r}
# Negative contribution means removing the cluster increased the mantel
# Towards the end, every part of the space is equivalent, no contribution is high
plot(sp_spread_sweep$alive_ratio, sp_spread_sweep$mantel_contribution)
plot(sp_spread_sweep$alive_ratio, sp_spread_sweep$mantel_contribution, ylim = c(0,1))
```



Plot iqr and some possible correlation plots

```{r}
save_pdf = FALSE
shape = "1col"

if (save_pdf){
  file_name = "202403-corr_iqr"
  prepare_figure_pdf(figure_folder, file_name, param_names, shape)
}

plot(sp_spread_sweep$iqr, sp_spread_sweep$false_positives, 
     pch=20, xlab = "Interquartile range", ylab = "False positives"
     )

if(save_pdf){
  dev.off()
}


if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-corr_mc_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 5,
    height = 4)
}

plot(sp_spread_sweep$mantel_contribution, sp_spread_sweep$false_positives,
     pch=20, xlab = "Mantel Contribution", ylab = "False positives"
     )

if(save_pdf){
  dev.off()
}


if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-corr_pvalue_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 5,
    height = 4)
}

plot(sp_spread_sweep$pval, sp_spread_sweep$false_positives,
     pch=20, xlab = "Cluster p-value", ylab = "False positives"
     )

if(save_pdf){
  dev.off()
}


if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-corr_mantel_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 5,
    height = 4)
}

plot(sp_spread_sweep$mantel, sp_spread_sweep$false_positives,
     pch=20, xlab = "Cluster Mantel coefficient", ylab = "False positives"
     )

if(save_pdf){
  dev.off()
}


if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-corr_res-m_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 5,
    height = 4)
}

plot(sp_spread_sweep$cropped_mantel, sp_spread_sweep$false_positives,
     pch=20, xlab = "Residual Mantel coefficient", ylab = "False positives"
     )

if(save_pdf){
  dev.off()
}


if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-corr_res-p_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 5,
    height = 4)
}

plot(sp_spread_sweep$cropped_pvalue, sp_spread_sweep$false_positives,
     pch=20, xlab = "Residual p-value", ylab = "False positives"
     )

if(save_pdf){
  dev.off()
}
```

Plot MC with inset

```{r}
save_pdf = FALSE

if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-mc_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 5,
    height = 4)
}

positives = subset(sp_spread_sweep, mantel_contribution>0)
inset = ggplot(sp_spread_sweep, aes(alive_ratio, mantel_contribution)) + geom_point(color="darkgray") +
  geom_point(data=positives, color="black") +
  # theme(axis.text.x = element_blank(), axis.ticks.x =element_blank() )+
  xlab("") +
  ylab("")

main_plot = ggplot(sp_spread_sweep, aes(alive_ratio, mantel_contribution)) + geom_point(color="darkgray") +
  geom_point(data = positives, color="black") +
  ylim(-0.2,1) + 
  xlab("Ratio of terraformed planets") +
  ylab("Mantel contribution")

#A viewport taking up a fraction of the plot area
vp <- viewport(width = 0.4, height = 0.4, x = 0.75, y = 0.7)

print(main_plot)
print(inset, vp = vp)

if(save_pdf){
  dev.off()
}

```



Plot selection space with MC threshold at 50%

```{r}

mantel_contribution_threshold = 0.5 # 0.1

if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-fig3_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 4,
    height = 4)
}

# make this a function

candidates_clusters = subset(sp_spread_sweep, iqr<iqr_threshold)
candidates_clusters = subset(candidates_clusters, mantel_contribution>mantel_contribution_threshold)

plot(sp_spread_sweep$iqr, sp_spread_sweep$mantel_contribution,
     xlab = "Interquartile range", ylab = "Mantel Contribution", pch=20, axes = FALSE,
     xlim = c(0,60), # ylim = c(5,35), 
     )
points(candidates_clusters$iqr, candidates_clusters$mantel_contribution,
       pch=20, col="red"
       )
axis(1)
axis(2)
abline(h=mantel_contribution_threshold)
abline(v=iqr_threshold)

if(save_pdf){
  dev.off()
}


```


Calculate detection performance metrics for individual clusters

```{r}

columns = c("true_positives", "false_positives", "true_negatives", "false_negatives",
            "accuracy", "sensitivity", "specificity")
sp_spread_sweep[columns] = 0

for (t in unique(sp_spread_sweep$step)){
  df_planets_t = subset(df_planets, step == t)
  sp_spread = subset(sp_spread_sweep, step == t)
  dbscan_cl = subset(dbscan_sweep, step == t)
  
  for (cid in unique(dbscan_cl$cluster_id)){
    detected = df_planets_t[dbscan_cl$cluster_id==cid,]
    perf = evaluate_detection(detected, df_planets_t)
    sp_spread_sweep[(sp_spread_sweep$step == t) & (sp_spread_sweep$cluster_id==cid), columns] = perf
  }
}

columns = c("true_positives", "false_positives", "true_negatives", "false_negatives",
            "accuracy", "sensitivity", "specificity")
```


Plot performance

```{r}
candidates_clusters = subset(sp_spread_sweep, iqr<iqr_threshold)
# candidates_clusters = subset(sp_spread_sweep, mantel_contribution>mantel_contribution_threshold)
candidates_clusters = subset(candidates_clusters, mantel_contribution>0)

plot(df_model$n_living_planets/1000, df_model$n_living_planets, type="l",
     main = "True positives", xlab = "Terraformed ratio"
     )
points(sp_spread_sweep$alive_ratio, sp_spread_sweep$true_positives)
points(candidates_clusters$alive_ratio, candidates_clusters$true_positives, col="red")
```


```{r}

candidates_clusters = subset(sp_spread_sweep, iqr<iqr_threshold)
plot(df_model$n_living_planets/1000, df_model$n_living_planets, type="l",
     main = "True positives (iqr selection)", xlab = "Terraformed ratio"
     )
points(sp_spread_sweep$alive_ratio, sp_spread_sweep$true_positives)
points(candidates_clusters$alive_ratio, candidates_clusters$true_positives, col="blue")

candidates_clusters = subset(sp_spread_sweep, mantel_contribution>=mantel_contribution_threshold)
plot(df_model$n_living_planets/1000, df_model$n_living_planets, type="l",
     main = "True positives (MC selection)", xlab = "Terraformed ratio"
     )
points(sp_spread_sweep$alive_ratio, sp_spread_sweep$true_positives)
points(candidates_clusters$alive_ratio, candidates_clusters$true_positives, col="blue")


candidates_clusters = subset(sp_spread_sweep, iqr<iqr_threshold)
plot(df_model$n_living_planets/1000, df_model$n_living_planets, type="l",
     main = "False positives (iqr selection)", xlab = "Terraformed ratio"
     )
points(sp_spread_sweep$alive_ratio, sp_spread_sweep$false_positives)
points(candidates_clusters$alive_ratio, candidates_clusters$false_positives, col="red")

candidates_clusters = subset(sp_spread_sweep, mantel_contribution>=mantel_contribution_threshold)
plot(df_model$n_living_planets/1000, df_model$n_living_planets, type="l",
     main = "False positives (MC selection)", xlab = "Terraformed ratio"
     )
points(sp_spread_sweep$alive_ratio, sp_spread_sweep$false_positives)
points(candidates_clusters$alive_ratio, candidates_clusters$false_positives, col="red")


# Both
candidates_clusters = subset(sp_spread_sweep, iqr<iqr_threshold)
candidates_clusters = subset(candidates_clusters, mantel_contribution>=mantel_contribution_threshold)

plot(df_model$n_living_planets/1000, df_model$n_living_planets, type="l",
     main = "True positives (iqr & MC selection)", xlab = "Terraformed ratio"
     )
points(sp_spread_sweep$alive_ratio, sp_spread_sweep$true_positives)
points(candidates_clusters$alive_ratio, candidates_clusters$true_positives, col="blue")

plot(df_model$n_living_planets/1000, df_model$n_living_planets, type="l",
     main = "False positives (iqr & MC selection)", xlab = "Terraformed ratio"
     )
points(sp_spread_sweep$alive_ratio, sp_spread_sweep$false_positives)
points(candidates_clusters$alive_ratio, candidates_clusters$false_positives, col="red")
```


Plot all performance measures for selected clusters for combined selection methods

```{r}
save_pdf = FALSE

mc_values = c(0,0.25,0.5)
if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  # figure_path = paste0(figure_folder, "/202403-performance_ratios_", param_names, ".pdf")
  figure_path = paste0(figure_folder, "/performance_ratios", ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 8,
    height = 3)
}

par(mfcol = c(1,3)) 

for (mantel_contribution_threshold in mc_values){

  candidates_clusters = subset(sp_spread_sweep, iqr<iqr_threshold)
  candidates_clusters = subset(candidates_clusters, mantel_contribution>mantel_contribution_threshold)

  ylabel = ""
  xlabel = ""
  if(mantel_contribution_threshold == 0){
    ylabel = "Ratio of planets"
    xlabel = "Ratio of terraformed planets"
  }
  
  plot(candidates_clusters$alive_ratio, candidates_clusters$true_positives/candidates_clusters$cluster_size, col="#117733", 
       ylim = c(0,1), xlim = c(0,1), lwd = 3, axes = FALSE, pch=20,
       xlab = xlabel, ylab = ylabel, 
       )
  axis(1)
  axis(2, pos=0)
  points(candidates_clusters$alive_ratio, candidates_clusters$false_positives/candidates_clusters$cluster_size, 
         col="#999933", lwd = 3, pch=20
         )
  points(candidates_clusters$alive_ratio, candidates_clusters$true_negatives/(1000-candidates_clusters$cluster_size), 
         col="#882255", lwd = 3, pch=20
         )
  points(candidates_clusters$alive_ratio, candidates_clusters$false_negatives/(1000-candidates_clusters$cluster_size), 
         col="#CC6677", lwd = 3, pch=20
         )
  if(mantel_contribution_threshold==0.5){
    legend(0.4, 0.7, legend=c("True positives", "False positives", "True negatives", "False negatives"),  
           fill = c("#117733","#999933", "#882255", "#CC6677")
    )
  }
  
}
  
  
if(save_pdf){
  dev.off()
}

```



Collate the data by MC threshold

```{r}
m = matrix(0, nrow = 0, ncol = 4)
for (mantel_contribution_threshold in mc_values){
  candidates_clusters = subset(sp_spread_sweep, iqr<iqr_threshold)
  candidates_clusters = subset(candidates_clusters, mantel_contribution>mantel_contribution_threshold)
  m = rbind(m, cbind(rep(mantel_contribution_threshold, length(candidates_clusters$cluster_id)), 
                     candidates_clusters[,c("alive_ratio", "sensitivity","specificity","accuracy")]))
}
perf_by_mc = data.frame(m)
colnames(perf_by_mc) = c("mc_threshold","alive_ratio", "sensitivity","specificity","accuracy")

perf_by_mc$mc_threshold = as.character(perf_by_mc$mc_threshold)
```


Performance scatterplots

```{r}
save_pdf = FALSE

if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-performance_metrics_",
                       mantel_contribution_threshold, "_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 8,
    height = 3)
}

par(mfcol = c(1,3)) 

sp <- ggplot(perf_by_mc, aes(alive_ratio, sensitivity)) + 
  geom_point(aes(color = mc_threshold)) +
  xlab("Ratio of terraformed planets") +
  ylim(0,1)+
  theme_minimal() +
  theme(legend.position = "top") +
  scale_color_brewer("Blues")
sp

sp <- ggplot(perf_by_mc, aes(alive_ratio, specificity)) + 
  geom_point(aes(color = mc_threshold)) +
  xlab("") +
  ylim(0,1)+
  theme_minimal()+
  theme(legend.position = "top") +
  scale_color_brewer("Blues")
sp

sp <- ggplot(perf_by_mc, aes(alive_ratio, accuracy)) + 
  geom_point(aes(color = mc_threshold)) +
  xlab("") +
  ylim(0,1)+
  theme_minimal()+
  theme(legend.position = "top") +
  scale_color_brewer("Blues")
sp

if(save_pdf){
  dev.off()
}
```


Performance boxplots

```{r}
short = as.data.table(perf_by_mc[, c( "mc_threshold", "sensitivity", "specificity", "accuracy")])
long = melt(short, id = c("mc_threshold"), value.name = "performance") 
long$mc_threshold <- as.factor(long$mc_threshold)
```


```{r}
# from https://www.datanovia.com/en/lessons/combine-multiple-ggplots-into-a-figure/
save_pdf = FALSE

if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  # figure_path = paste0(figure_folder, "/202403-boxplots_", param_names, ".pdf")
  figure_path = paste0(figure_folder, "boxplots", ".pdf")
  
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 8,
    height = 3)
}

bp <- ggplot(long, aes(mc_threshold, performance)) + 
  geom_boxplot(aes(fill = mc_threshold), position = position_dodge(0.9)) +
  scale_fill_brewer(palette = "Blues", name = "MC threshold") +
  theme_minimal()+
  theme(legend.position = "top") +
  xlab("Mantel Contribution Threshold") +
  ylab("")
bp + facet_grid(cols = vars(variable))

if(save_pdf){
  dev.off()
}
```



Finally, plot an example for 1 time step

```{r}
save_pdf = FALSE


# first acceptable detection for MC threshold = 0.25
t = 67

df_planets_t = subset(df_planets, step == t)
dbscan_cl = subset(dbscan_sweep, step == t)
terraformed = subset(df_planets_t, alive == alive_value)
detected = df_planets_t[dbscan_cl$cluster_id==cid,]


if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  #figure_path = paste0(figure_folder, "/202403-pca_", param_names, ".pdf")
  figure_path = paste0(figure_folder, "pca", ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 5,
    height = 4)
}

# Plot pca
compositions = df_planets_t[,select_cols]
pca_res <- prcomp(compositions, scale. = TRUE)
autoplot(pca_res, col=dbscan_cl$cluster+1, pch=20)

if(save_pdf){
  dev.off()
}

if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-detection_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 5,
    height = 4)
}

# plot terraformed planets
plot(df_planets_t$x, df_planets_t$y,
     pch=20, col="grey", xlab = "X",  ylab = "Y"
     )
points(terraformed$x, terraformed$y, col="blue", pch=20)
points(detected$x, detected$y, col="red")
  
if(save_pdf){
  dev.off()
}


```
Plot some elbows

```{r}
save_pdf = FALSE

if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-elbow_detection_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 5,
    height = 4)
}


# early elbow low sensitivity
t = 105
df_planets_t = subset(df_planets, step == t)
compositions = df_planets_t[,select_cols]
comp_distances = full(distance(compositions))

par(mfcol = c(2,2)) 
epsilon_dbscan(comp_distances, db_minPts, total_n_planets, plot_elbow = TRUE, my_sensitivity = elbow_sens)

# early elbow high sensitivity
d = epsilon_dbscan(comp_distances, db_minPts, total_n_planets, plot_elbow = TRUE, my_sensitivity = 4)

# late elbow low sensitivity
t = 205
df_planets_t = subset(df_planets, step == t)
compositions = df_planets_t[,select_cols]
comp_distances = full(distance(compositions))
epsilon_dbscan(comp_distances, db_minPts, total_n_planets, plot_elbow = TRUE, my_sensitivity = elbow_sens)

# late elbow high sensitivity
epsilon_dbscan(comp_distances, db_minPts, total_n_planets, plot_elbow = TRUE, my_sensitivity = 10)

if(save_pdf){
  dev.off()
}


```



Plot the iqr or number of planets and then use contribution as color

```{r}
positives = subset(sp_spread_sweep, mantel_contribution>0)
sc = scale_color_gradient(low = "blue", high = "red", 
                          limits = c(0,1),
                          name = "MC")

ggplot(positives, aes(alive_ratio, iqr)) +
  geom_point(aes(color = mantel_contribution)) + sc +
  geom_hline(yintercept = iqr_threshold, linetype="dashed") +
  xlab("Percentage of terraformed planets") +
  ylab("Spatial IQR")


ggplot(positives, aes(alive_ratio, cluster_size)) +
  geom_point(aes(color = mantel_contribution)) + sc +
  xlab("Percentage of terraformed planets") +
  ylab("Number of planets in cluster")

ggplot(positives, aes(alive_ratio, pval)) +
  geom_point(aes(color = mantel_contribution)) + sc +
  xlab("Percentage of terraformed planets") +
  ylab("Cluster p value")

ggplot(positives, aes(alive_ratio, pval)) +
  geom_point(aes(color = mantel_contribution)) + sc +
  xlab("Percentage of terraformed planets") +
  ylab("Cluster p value") +
  ylim(0,0.05)

```

```{r}
save_pdf = FALSE

if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-iqr_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 5,
    height = 4)
}

ggplot(positives, aes(alive_ratio, iqr)) +
  geom_point(aes(color = mantel_contribution)) + sc +
  geom_hline(yintercept = iqr_threshold, linetype="dashed") +
  theme(legend.position = c(0.93, 0.80)) +
  xlab("Ratio of terraformed planets") +
  ylab("Spatial IQR")

if(save_pdf){
  dev.off()
}
```
