---
title: "Figure1"
author: "Lana"
date: '2023-06-22'
output: html_document
---

This file plots the mantel heatmaps.

Libraries

```{r}
library("ggplot2")     
source("Utilities.R")
library("beepr")
library("patchwork")
```


Read compressed files (.gz data) for radius selection function

```{r}
setwd("../analysis")
base_path = "../scripts/example_data/"
sub_folder = "2024-03-12-175828/"
figure_folder = paste0(base_path, sub_folder, "figures/")
dir.create(figure_folder, showWarnings = FALSE)


data = read_sim_data(base_path, sub_folder)
df_model_sweep = data[[4]]
df_parameters = data[[1]]


# Other parameters
space_dim = 3
comp_dim = df_parameters$compsize
# number of permutations for mantel calculation
n_perm = 100
comp_names = paste("comp_",c(1:comp_dim), sep="")
selection_func = "r"
```


Remove some of the redundant r values
```{r}
# 12, 11 and 10 have ~ the same values as 20; remove all and keep 20
df_parameters$r = c(9, 20, 50, 174)
df_model_sweep = subset(df_model_sweep, r %in% df_parameters$r)
```


## Parameter selection for heatmaps

Choose a parameter subdivision to create heatmaps
Plot different subdivisions of the model for visual examination

```{r}
divisions = c("nool", "n_idxs_to_keep_from_destination","r", "mutation_rate")
plot_divisions(df_model_sweep, df_parameters, divisions)
```


Choose a division, eg ool, and plot the other parameters by color again
If it looks good, this is how we will divide the data for the heatmaps.

```{r}
# The data is subset according to the ool values, then color-coded by the other parameters.
plot_subdivision("nool", df_model_sweep, df_parameters, divisions)
```


## Curve selection: Plot all curves to pick representative ones

(Skip to next cell if the focus scenarii have already been chosen)

```{r}
plot_runs_individually(df_parameters, df_model_sweep, selection_func)
```


Those are our chosen curves.
These curves will be highlighted on the heatmaps

```{r}

# diff and id must not be the same for 2 choices
# curvetype: 0 is flat, 1 is growth, 2 is inverted U
cols = c( "ool", "select_value", "mutation", "id", "curve_type")
chosen_dfs = data.frame(matrix(0, nrow = 0, ncol = length(cols)-1))
selection = c()

# ool r mut id
# ool 10
# growth
selection = c(selection, "r")
chosen = c(10, 20, 0, 0, 1)
chosen_dfs = rbind(chosen_dfs, chosen)
# no u
# flat
selection = c(selection, "r")
chosen = c(10, 9, 0, 5, 0)
chosen_dfs = rbind(chosen_dfs, chosen)

# ool 1
# growth
selection = c(selection, "r")
chosen = c(1, 20, 0, 2, 1)
chosen_dfs = rbind(chosen_dfs, chosen)
# inverted U
selection = c(selection, "r")
chosen = c(1, 174, 0, 0, 2)
chosen_dfs = rbind(chosen_dfs, chosen)
# flat
selection = c(selection, "r")
chosen = c(1, 174, 0, 5, 0)
chosen_dfs = rbind(chosen_dfs, chosen)

names(chosen_dfs) = cols
chosen_dfs$selection = selection
```


Those are our chosen curves for startingmantel=high
10 0.8 0 5 (flat)
10, 0.8, 0, 0, 0 (curved)
1 0.8 0 0 (short), 
1 3.2 0.1 0 (down)


# Now make the actual figures

Heatmaps with highlights and labels, for starting with low mantel

```{r}
# overlay on heatmap
# https://stackoverflow.com/questions/64309620/strikethrough-overlay-on-specific-cells-of-a-heatmap-in-ggplot2

# colorblind friendly colors
# http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
# flat, growth, inverted u
colors = c("#D55E00","#F0E442", "#009E73")

ggp_heat_r_1 = heatmap_with_highlight(df_model_sweep, selection_func, chosen_dfs, df_parameters, my_n_ool=1, my_colors=colors, legend = FALSE)
ggp_heat_r_1
ggp_heat_r_10 = heatmap_with_highlight(df_model_sweep, selection_func, chosen_dfs, df_parameters, my_n_ool=10, my_colors=colors, legend = FALSE)
ggp_heat_r_10
```

Draw all curves from selected cells in same graph

```{r}
plot_selected_r = plot_selected_mantels(df_model_sweep, chosen_dfs, "r")
plot_selected_r
```



Finally, put all figures in 1 plot

```{r}
save_pdf = TRUE

file_path = paste0(figure_folder, "/202403-heatmap.pdf")
make_heatmap_figure("r", ggp_heat_r_1, ggp_heat_r_10, plot_selected_r, my_save = save_pdf, my_filename = file_path)
```
