---
title: "202403-fig1"
author: "Lana"
date: "2024-03-07"
output: html_document
---

This file plots one simulation run's mantel and pvalue.

Load libraries

```{r}
source("Utilities.R")
```


Set folder and type of selection function here

```{r}
base_path = "../scripts/example_data/"
sub_folder = "2024-03-12-175828/"
selection_func = "r"

figure_folder = paste0(base_path, sub_folder, "figures/")
dir.create(figure_folder, showWarnings = FALSE)
```


Read data

```{r}
data = read_sim_data(base_path, sub_folder)

df_model_sweep = data[[4]]
df_parameters = data[[1]]
df_planets_sweep = data[[2]]
df_life_sweep = data[[3]]
diffs = df_parameters$r
param_order = "ool r mut id"
selection_func = "r"

# Remove some of the redundant r values
# 12, 11 and 10 have ~ the same values as 20; remove all and keep 20
df_parameters$r = c(9, 20, 50, 174)
df_model_sweep = subset(df_model_sweep, r %in% df_parameters$r)
```


Run this when selecting one run for further analysis. If the run is already selected go to the next cell.
For plots divided by parameters, see "Figure1.Rmd" file.
Plot all the mantels together, then one by one in radius selection function.

```{r}
plot(df_model_sweep$n_living_planets/1000, df_model_sweep$mantel_corr_coeff, main=sub_folder)

ids = df_parameters$n_idxs_to_keep_from_destination
mutations = df_parameters$mutation_rate
ools = df_parameters$nool

for (n_ool in ools){
  df_model_ool = subset(df_model_sweep,  nool==n_ool)
  for (n_diff in diffs){
    df_model_diff = subset(df_model_ool,  df_model_ool[,selection_func]==n_diff)
      for (n_mutation in mutations){
        df_model_mut = subset(df_model_diff,  mutation_rate==n_mutation)
        for (n_id in ids){
          df_model_id = subset(df_model_mut,  n_idxs_to_keep_from_destination==n_id)
          title = paste(param_order, n_ool, n_diff, n_mutation, n_id)
          
          plot(df_model_id$n_living_planets/1000, df_model_id$mantel_corr_coeff,
               ylim = c(0,0.5), xlim = c(0,1),
               main = title
               )
      }
    }
  }
}
```


Subset 1 time step of 1 run.

```{r}
n_ool = 1
p_r = 20
p_mutation = 0
p_id = 1

result = select_parameter_set(df_planets_sweep, df_life_sweep, p_id, p_r, p_mutation, selection_func)

df_planets = result[[1]]
df_lifes = result[[2]]
df_model = result[[3]]
```


Plot Fig 1 with 2 y axes

```{r}
save_pdf = TRUE
if (save_pdf){
  param_names = paste0("pool_", n_ool, "-pid_", p_id, "-pr_", p_r, "-mutr_", p_mutation)
  figure_path = paste0(figure_folder, "/202403-one_mantel_", param_names, ".pdf")
  print(paste("Saving figure at", figure_path))
  pdf(file = figure_path,   
    width = 5,
    height = 4)
}

par(mar = c(5, 4, 4, 4) + 0.3)              # Additional space for second y-axis
plot(df_model$n_living_planets/1000, df_model$mantel_corr_coeff,
     ylim = c(0,0.2), xlim = c(0,1),
     ylab = "Mantel coefficient", xlab = "Ratio of terraformed planets", axes = FALSE,
     col = "black", type = "l")
axis(1, pos = 0)
axis(2, pos = 0)
par(new = TRUE)                             # Add new plot
plot(df_model$n_living_planets/1000, df_model$mantel_p_value,
     ylim = c(0,0.4), xlim = c(0,1), col = "red",  type = "l",             # Create second plot without axes
     axes = FALSE, xlab = "", ylab = "")
axis(side = 4, at = pretty(c(0,0.4)), pos = 1,
     col="red", col.ticks="red", col.axis="red")      # Add second axis
mtext("p-value", side = 4, line = 3, col="red")    

if(save_pdf){
  dev.off()
}

```

